---
title: "WIC Transaction Data"
output: html_notebook
---

1) Import Data from csv, skipping the rows which are in the header of the document, into R dataframes
2) Combine all dataframes into a single dataframe
3) Report out the number of unique families
4) Report out the average spending per family
5) Report out the average number of vendors a family uses

#getting the directory

```{r}
#print(getwd())

```
#library used

```{r}
library(dplyr)
library(sqldf)
library(ggplot2)
library(modeest)

```


#https://stackoverflow.com/questions/39110755/skip-specific-rows-using-read-csv-in-r

#Reading the first file for Guilford_County_Redemption_Week1, loading the data by skipping the documentation row from row 1 to row 14

```{r}
headers = read.csv("Guilford_County_Redemption_Week1.csv", skip = 14, header = F, nrows = 1, as.is = T)
Redemption_df1 = read.csv("Guilford_County_Redemption_Week1.csv", skip = 15, header = F)
colnames(Redemption_df1)= headers
```


#Reading the Second file for Guilford_County_Redemption_Week2, loading the data by skipping the documentation row from row 1 to row 14


```{r}
headers2 = read.csv("Guilford_County_Redemption_Week2.csv", skip = 14, header = F, nrows = 1, as.is = T)
Redemption_df2 = read.csv("Guilford_County_Redemption_Week2.csv", skip = 15, header = F)
colnames(Redemption_df2)= headers

```


#Reading the Third file for Guilford_County_Redemption_Week,3 loading the data by skipping the documentation row from row 1 to row 14


```{r}
headers3 = read.csv("Guilford_County_Redemption_Week3.csv", skip = 14, header = F, nrows = 1, as.is = T)
Redemption_df3 = read.csv("Guilford_County_Redemption_Week3.csv", skip = 15, header = F)
colnames(Redemption_df3)= headers

```


#2) combining all the 3 dataframes together

```{r}
WIC_Redemption_df<-rbind(Redemption_df1,Redemption_df2,Redemption_df3)

#write.csv(WIC_Redemption_df,"WIC_Redemption_df.csv")

```

#3) Report out the number of unique families
#https://dplyr.tidyverse.org/reference/distinct.html

#6617 unique families

# count of unique families

```{r}
length(unique(WIC_Redemption_df$`WIC Family ID`))

```
#reporting out unique families 

```{r}

Unique_WIC_family_df<-distinct(WIC_Redemption_df, WIC_Redemption_df$'WIC Family ID', .keep_all = TRUE)

```

#Renaming columns for easy identification

```{r}
colnames(WIC_Redemption_df)[23] <- "Requested_Amount"
colnames(WIC_Redemption_df)[13] <- "WIC_Family_Id"
colnames(WIC_Redemption_df)[25] <- "Authorization_Number"
colnames(WIC_Redemption_df)[27] <- "Vendor_Number"
colnames(WIC_Redemption_df)[2] <- "Report_End_Date"
colnames(WIC_Redemption_df)[1] <- "Report_Begin_Date"
colnames(WIC_Redemption_df)[26] <- "Vendor_Name"

```



#Renaming columns for easy identification

```{r}
colnames(Unique_WIC_family_df)[23] <- "Requested_Amount"
colnames(Unique_WIC_family_df)[13] <- "WIC_Family_Id"
colnames(Unique_WIC_family_df)[25] <- "Authorization_Number"

```



#4) Report out the average spending per family per trip


```{r}

Average_spending_df<-WIC_Redemption_df %>%
   group_by(WIC_Family_Id, Authorization_Number) %>% 
   summarise(Avg_requested_amount = mean(Requested_Amount),No_of_items=n())
```


#5) Report out the average (mean and median) number of vendors a family uses


#count of unique vendors

```{r}
length(unique(WIC_Redemption_df$Vendor_Number))

```

#count number of vendors per item by grouping family and vendor number


```{r}
   count_number_of_item_per_vendor<-WIC_Redemption_df %>%
   group_by(WIC_Family_Id, Vendor_Number) %>% 
   summarise(No_of_item_per_vendor=n())
```
# vendor number occurance for each families 

```{r}

unique_vendor_number_count_for_families=count_number_of_item_per_vendor %>% group_by(Vendor_Number) %>% tally()


```

#finding frequency or count for vendor used/occurance for each families

```{r}
Vendor_number_freq<-sqldf("SELECT WIC_Family_Id, Vendor_Number, COUNT(*) as Vendor_number_freq
       FROM count_number_of_item_per_vendor
       GROUP BY WIC_Family_Id")
```


#finding the mean and median for each familiy for vendor used


```{r}

Avg_Vendor_freq<-sqldf("SELECT WIC_Family_Id, Vendor_Number, Vendor_number_freq, avg(Vendor_number_freq), median (Vendor_number_freq)
       FROM Vendor_number_freq
       GROUP BY WIC_Family_Id,Vendor_Number ")

Avg_Vendor_freq$mean_vendor <- Avg_Vendor_freq$Vendor_number_freq
Avg_Vendor_freq$median_vendor <- Avg_Vendor_freq$Vendor_number_freq

```


#grouping using week for each families

#Creating a new column week to identify if purchase was made in week1 or week2 or week3 based on date range(report start or report end date)

```{r}
Week_distribuion<- WIC_Redemption_df
Week_distribuion$week <- ifelse (WIC_Redemption_df$Report_Begin_Date == '2/17/2021' & WIC_Redemption_df$Report_End_Date == '2/23/2021','Week1',ifelse(WIC_Redemption_df$Report_Begin_Date == '2/24/2021' & WIC_Redemption_df$Report_End_Date == '3/2/2021','Week2','week3' ))



```


#Average spending amount grouped using family id, autorization number (each trip) and week (week 1, week 2 and week3)


```{r}
   Average_spending_week_df<-Week_distribuion %>%
   group_by(WIC_Family_Id, Authorization_Number, week) %>% 
   summarise(Avg_requested_amount = Requested_Amount,No_of_items=n())
```


#finding mean for each week

```{r}
Mean_weekly<-group_by(Average_spending_week_df, week) %>%  summarise(m=mean(Avg_requested_amount))
colnames(Mean_weekly)[2] <- "Mean"
```

#finding median for each week

```{r}

Median_weekly<-group_by(Average_spending_week_df, week) %>%  summarise(m= median (Avg_requested_amount))
colnames(Median_weekly)[2] <- "Median"
colnames(Median_weekly)[1] <- "week1"

```


#combining mean and median in one dataframe

```{r}
Mean_meadian_weekly<-cbind(Mean_weekly,Median_weekly)
Mean_meadian_weekly<- sqldf('SELECT week,Mean,Median FROM Mean_meadian_weekly' )
```

#Mode function

```{r}
#my_mode <- function(x) {                      
 # unique_x <- unique(x)
 # tabulate_x <- tabulate(match(x, unique_x))
 # unique_x[tabulate_x == max(tabulate_x)]
#}
```
#filter week data

```{r}
#filter_week_data<- filter(Average_spending_week_df,Average_spending_week_df$week == 'week1' )
```

#check mode function

```{r}
#my_mode(Average_spending_week_df$Avg_requested_amount)
```


#creating a dataframe with family id, vendor name and vendor number by grouoing on family id and vendor number

```{r}

family_with_vendor_name_df <- sqldf("SELECT WIC_Family_Id, Vendor_Number, Vendor_Name,Authorization_Number, COUNT(*) as No_of_item_per_vendor
       FROM WIC_Redemption_df
       GROUP BY WIC_Family_Id,Authorization_Number,Vendor_Number")


```

#from family_with_vendor_name_df dataframe counting the occurance of vendor name for a family per purchase trip by grouping on family id

```{r}
Vendor_names_freq_with_family_id<-sqldf("SELECT WIC_Family_Id, Vendor_Number, Vendor_Name, COUNT(*) as Vendor_number_freq
       FROM family_with_vendor_name_df
       GROUP BY WIC_Family_Id")
```


#Data visualizations

#Reporting average spending for each family

#shows the average spending for each family for each week

```{r}



ggplot(Mean_meadian_weekly, aes(x=week, y=Mean , Group = interaction(Median) )) + 
    geom_point(
      color="orange",
        fill="#69b3a2",
        shape=21,
        alpha=0.5,
        size=6,
        stroke = 2
    ) +
    ylim(0, 10)



ggplot(Mean_meadian_weekly, aes(week)) + geom_point(aes(y=Mean), color="red") + geom_point(aes(y=Median), color = "green")  +
    ylim(0, 10)


```


```{r}
ggplot(Average_spending_week_df, aes(x=week, y=Avg_requested_amount  )) + 
    geom_point(
      color="orange",
        fill="#69b3a2",
        shape=21,
        alpha=0.5,
        size=6,
        stroke = 2
    ) +
    ylim(0, 500)
```

#filtering certain vendors from Vendor_names_freq_with_family_id dataframe

```{r}
filter_vendor<- filter(Vendor_names_freq_with_family_id,Vendor_Number  == 14 | Vendor_Number  == 40 | Vendor_Number== 1116) 
```

#checking the data type for each column for filter_vendor dataframe

```{r}
str(filter_vendor)
```

#bar plot showing list of vendor name and family id and their vendor frequency

```{r}
  barplot(height= filter_vendor$Vendor_number_freq , names= filter_vendor$Vendor_Name, 
        col=rgb(0.8,0.1,0.1,0.6),
        xlab="Vendor Name", 
        ylab="Vendor frquency count", 
        main="Vendor distribution", 
        ylim=c(0,10)
        )
  barplot(height= filter_vendor$Vendor_number_freq , names= filter_vendor$WIC_Family_Id, 
        col=rgb(0.8,0.1,0.1,0.6),
        xlab="Familes", 
        ylab="Vendor frequency count", 
        main="Vendor distribution", 
        ylim=c(0,10)
        )
```


#plt showing no of item puchased from vendor

```{r}
ggplot(family_with_vendor_name_df, aes(x=Vendor_Name, y=No_of_item_per_vendor)) + 
    geom_point()
```

