---
title: "WIC Transaction Data- Merged"
output: html_notebook
---

#library used

```{r}
library(dplyr)
library(sqldf)

```
#load data
```{r}
Redemption_data_df <- read.csv("MergedData.csv")
```

#renaming columns for easy identification

```{r}
colnames(Redemption_data_df)[20] <- "Requested_Amount"
colnames(Redemption_data_df)[2] <- "Report_End_Date"
colnames(Redemption_data_df)[1] <- "Report_Begin_Date"
colnames(Redemption_data_df)[9] <- "Category_Code"
colnames(Redemption_data_df)[10] <- "Category_Name"
colnames(Redemption_data_df)[22] <- "Vendor_Name_x"
colnames(Redemption_data_df)[23] <- "Vendor_Number"
colnames(Redemption_data_df)[34] <- "Vendor_Name_y"
colnames(Redemption_data_df)[35] <- "Address"
colnames(Redemption_data_df)[36] <- "City"
colnames(Redemption_data_df)[37] <- "Zipcode"


```


#filter out positive values from Requested_Amount field

```{r}
Redemption_data_df1 <- Redemption_data_df %>%
            filter(Requested_Amount > 0)
```

#count of na values in Redemption_data dataframe

```{r}
sum(is.na(Redemption_data_df1))
```

#remove na values from Redemption_data dataframe

```{r}
Redemption_data_df1 <- na.omit(Redemption_data_df1)

```

#Select few columns from Redemption data 

```{r}
Redemption_data_df2<-sqldf("SELECT Report_Begin_Date,Report_End_Date,WIC_Fam_ID_Hash, Auth_Num_Hash,Avg(Requested_Amount) as Avg_Requested_Amount,Category_Code,Category_Name,Vendor_Name_x,Vendor_Number,Vendor_Name_y, Address, City, Zipcode
       FROM Redemption_data_df1
       GROUP BY WIC_Fam_ID_Hash,Auth_Num_Hash")
```


```{r}
Redemption_data_df2$Avg_Requested_Amount<-format(round(Redemption_data_df2$Avg_Requested_Amount, 2), nsmall = 2) 
```



#1) Report out the number of unique families

```{r}
length(unique(Redemption_data_df$WIC_Fam_ID_Hash))
length(unique(Redemption_data_df$Auth_Num_Hash))

```

#reporting out unique families 

```{r}
Unique_WIC_family_df<-distinct(Redemption_data_df, Redemption_data_df$WIC_Fam_ID_Hash, .keep_all = TRUE)
```

#Renaming column for easy identification

```{r}
colnames(Unique_WIC_family_df)[20] <- "Requested_Amount"

```
#Renaming column for easy identification

```{r}
colnames(Redemption_data_df)[20] <- "Requested_Amount"
colnames(Redemption_data_df)[2] <- "Report_End_Date"
colnames(Redemption_data_df)[1] <- "Report_Begin_Date"
```


#1)Report out the average spending per family per trip


```{r}
Average_spending_df<-Redemption_data_df %>%
   group_by(WIC_Fam_ID_Hash, Auth_Num_Hash) %>% 
   summarise(Avg_requested_amount = mean(Requested_Amount),  No_of_items=n())
```
#grouping data using week for each families

```{r}

Week_distribution<- Redemption_data_df
Week_distribution$Week <- ifelse (Redemption_data_df$Report_Begin_Date == '2/17/2021' & Redemption_data_df$Report_End_Date == '2/23/2021','Week1',ifelse(Redemption_data_df$Report_Begin_Date == '2/24/2021' & Redemption_data_df$Report_End_Date == '3/2/2021','Week2','week3' ))

```

#rounding the data to 2 decimal places

```{r}
Week_distribution$Requested_Amount<-format(round(Week_distribution$Requested_Amount, 2), nsmall = 2)     
```
#Week_distribution dataframe grouped by family id, auth number and week

```{r}
Average_spending_week_df1<-sqldf("SELECT WIC_Fam_ID_Hash, Auth_Num_Hash,Requested_Amount, AVG(Requested_Amount) as Avg_Requested_amount,Week
       FROM Week_distribution
       GROUP BY WIC_Fam_ID_Hash,Auth_Num_Hash,Week")


```

#rounding the data to 2 decimal places

```{r}

Average_spending_week_df1$Avg_Requested_amount<-format(round(Average_spending_week_df1$Avg_Requested_amount, 2), nsmall = 2) 
 
```

#removing na 

```{r}
Average_spending_week_df1<- na.omit(Average_spending_week_df1)
```

#find mean of data

```{r}
#tapply(Average_spending_week_df1$Requested_Amount,Average_spending_week_df1$Week, mean)
```



```{r}

```

