---
title: "WIC Transaction Data"
output: html_notebook
---

1) Import Data from csv, skipping the rows which are in the header of the document, into R dataframes
2) Combine all dataframes into a single dataframe
3) Report out the number of unique families
4) Report out the average spending per family
5) Report out the average number of vendors a family uses

#getting the directory

```{r}
print(getwd())

```
#https://stackoverflow.com/questions/39110755/skip-specific-rows-using-read-csv-in-r

#Reading the first file for Guilford_County_Redemption_Week1, loading the data by skipping the documentation row from row 1 to row 14

```{r}
headers = read.csv("Guilford_County_Redemption_Week1.csv", skip = 14, header = F, nrows = 1, as.is = T)
Redemption_df1 = read.csv("Guilford_County_Redemption_Week1.csv", skip = 15, header = F)
colnames(Redemption_df1)= headers
```


#Reading the Second file for Guilford_County_Redemption_Week2, loading the data by skipping the documentation row from row 1 to row 14


```{r}
headers2 = read.csv("Guilford_County_Redemption_Week2.csv", skip = 14, header = F, nrows = 1, as.is = T)
Redemption_df2 = read.csv("Guilford_County_Redemption_Week2.csv", skip = 15, header = F)
colnames(Redemption_df2)= headers

```


#Reading the Third file for Guilford_County_Redemption_Week,3 loading the data by skipping the documentation row from row 1 to row 14


```{r}
headers3 = read.csv("Guilford_County_Redemption_Week3.csv", skip = 14, header = F, nrows = 1, as.is = T)
Redemption_df3 = read.csv("Guilford_County_Redemption_Week3.csv", skip = 15, header = F)
colnames(Redemption_df3)= headers

```


#combining all the 3 dataframes together

```{r}
WIC_Redemption_df<-rbind(Redemption_df1,Redemption_df2,Redemption_df3)
```

#3) Report out the number of unique families
#https://dplyr.tidyverse.org/reference/distinct.html

#6617 unique families

# count of unique families

```{r}
length(unique(WIC_Redemption_df$`WIC Family ID`))

```
#reporting out unique families 

```{r}

Unique_WIC_family_df<-distinct(WIC_Redemption_df, WIC_Redemption_df$'WIC Family ID', .keep_all = TRUE)

```



#4) Report out the average spending per family
# WIC family id with mean spending(Requested Amount) only these columns?

#5) Report out the average number of vendors a family uses
#for each family unique vendors
#find unique vendors, unique family ?
#WIC family id with mean vendors ? or count of vendors (Vendor id) 
#only 2 columns vendors and WIC family id? 



